On exponential sums arising from the classical doubling method

### 2025-03-28 main.tex
Minors

@@

Let $$w_n = \begin{pmatrix}
	& & & 1\\
	& & 1\\
	& \iddots\\
	1
\end{pmatrix} \in \GL_n\left(\finiteField\right)$$ be the longest Weyl element.
\begin{enumerate}
	\item (Special orthogonal groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = 1$. It follows by definition that every element $g \in \IsometryGroup\left(\hermitianSpace\right)$ satisfies $\left(\det g\right)^2 = 1$.  In this case, we obtain the connected components of the identity by adding a restriction on the determinant: We have that $G$ consists of all elements $g \in \IsometryGroup\left(\hermitianSpace\right)$ with $\det g = 1$. There are two cases here.
	\begin{enumerate}
		\item Split case: in this case, we can write $$\hermitianSpace = \begin{dcases}
		\xIsotropic \oplus \yIsotropic & \dim_{\finiteField} \hermitianSpace = 2n\\
		\xIsotropic \oplus \finiteField v \oplus \yIsotropic & \dim_{\finiteField}\hermitianSpace = 2n + 1
		\end{dcases}$$ where $\xIsotropic$ and $\yIsotropic$ are totally isotropic (that is, for every $x \in \xIsotropic$, $\innerproduct{x}{x} = 0$ and similarly for $\yIsotropic$) and these spaces are in duality with respect to $\innerproduct{\cdot}{\cdot}$, i.e., the map $\yIsotropic \to \Hom_{\finiteField}\left(\xIsotropic, \finiteField\right)$ given by $y \mapsto \left(x \mapsto \innerproduct{x}{y}\right)$ is an isomorphism. In the odd dimensional case, $v \in \hermitianSpace$ is an element orthogonal to $\xIsotropic$ and $\yIsotropic$ that satisfies $\innerproduct{v}{v} \ne 0$.

@@

	\begin{theorem}
		A maximal rational torus $\algebraicGroup{T}$ of $\algebraicGroup{G} = \algebraicGroup{\UnitaryGroup}_{n}, \algebraicGroup{\Sp}_{2n}, \algebraicGroup{\SO}_{2n + 1}, \algebraicGroup{\SO}_{2n}^{\pm}$ is isomorphic to 	$$\algebraicGroup{T} \cong \restrictionOfScalars{\finiteFieldExtension{\lambda^{+}}}{\finiteField}{\multiplcativeScheme} \times \restrictionOfScalars{\finiteFieldExtension{\lambda^{-}}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1},$$
		where $\left(\lambda^+, \lambda^-\right)$ is a pair of partitions such that $\sizeof{\lambda^+} + \sizeof{\lambda^-} = n$, where in the unitary group case we require that $\lambda^+$ consists only of even parts and that $\lambda^-$ consists only of odd parts. We have that $$\algebraicGroup{T}^{\Frobenius} \cong \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \times \NormOneGroup{2 \lambda^{-}}.$$ The embedding $\algebraicGroup{T}^{\Frobenius} \to \algebraicGroup{G}^{\Frobenius}$ is of the form $$\left(a, t\right) \mapsto
		\begin{pmatrix}
			\iota_+\left(a\right)\\
			& \iota_{-}\left(t\right) &\\
			 & & \involution{\iota_+^{\ast}\left(a\right)}
		\end{pmatrix},$$
		where $\iota_+ \colon \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \to \GL_{\frac{\sizeof{\lambda^+}}{\grpIndex{\quadraticExtension}{\finiteField}}}\left(\quadraticExtension\right)$ and $\iota_- \colon \NormOneGroup{2 \lambda^{-}} \to G'$ are embeddings. Here, $G'$ is a group of the same type as $\algebraicGroup{G}^{\Frobenius}$, with $n$ replaced by $\sizeof{\lambda^{-}}$.

@@

	We summarize these cases in the following theorem.
	\begin{theorem}
	A maximal rational torus $\algebraicGroup{T}$ of $\algebraicGroup{G} =  \algebraicGroup{\GSp}_{2n}, \algebraicGroup{\GSO}_{2n}^{\pm}$ is isomorphic to \begin{align*}
				^{L^{-1}w}\algebraicGroup{T}_0 \cong& \left\{\left(x,y,z\right) \in  \restrictionOfScalars{\finiteFieldExtension{\lambda^+}}{\finiteField}{\multiplcativeScheme} \times \restrictionOfScalars{\finiteFieldExtension{2\lambda^-}}{\finiteField}{\multiplcativeScheme}  \times \multiplcativeScheme \mid \FieldNorm{2\lambda^-}{\lambda^-}(y) = z \right\}.
	\end{align*}
	where $\left(\lambda^+, \lambda^-\right)$ is a pair of partitions such that $\sizeof{\lambda^+} + \sizeof{\lambda^-} = n$. We have that $$\algebraicGroup{T}^{\Frobenius} \cong \left\{\left(x, y, z\right) \in \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \times \multiplicativegroup{\finiteFieldExtension{2 \lambda^-}} \times \multiplicativegroup{\finiteField} \mid \FieldNorm{2 \lambda^{-}}{\lambda^{-}}\left(t_i\right) = z \right\}.$$ The embedding $\algebraicGroup{T}^{\Frobenius} \to \algebraicGroup{G}^{\Frobenius}$ is of the form $$\left(x, y, z\right) \mapsto
	\begin{pmatrix}
		z \cdot \iota_+ \left(x\right)\\
		& \iota_{-}\left(y\right) &\\
		& & \iota_+^{\ast}\left(x\right)
	\end{pmatrix},$$
	where $\iota_+ \colon \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \to \GL_{\sizeof{\lambda^+}}\left(\finiteField\right)$ and $$\iota_- \colon \left\{ \left(y, z\right)\in \multiplicativegroup{\finiteFieldExtension{2 \lambda^-}} \times \multiplicativegroup{\finiteField} \mid \FieldNorm{2 \lambda^{-}}{\lambda^{-}}\left(y\right) = z \text{ for all } i \right\} \to G'$$ are embeddings. Here, $G'$ is a group of the same type as $\algebraicGroup{G}^{\Frobenius}$, with $n$ replaced by $\sizeof{\lambda^{-}}$.
\end{theorem}

>>>>

\elad{Need to define $2 \lambda$ and $\FieldNorm{2\lambda}{\lambda}$ and $\NormOneGroup{2 \lambda}$}

@@

Let $$w_n = \begin{pmatrix}
	& & & 1\\
	& & 1\\
	& \iddots\\
	1
\end{pmatrix} \in \GL_n\left(\finiteField\right)$$ be the longest Weyl element.
\calvin{we should standardize an order between Unitary, Special Orthogonal Symplectic and try to kinda stick with it throughout.}
\begin{enumerate}
	\item (Special orthogonal groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = 1$. It follows by definition that every element $g \in \IsometryGroup\left(\hermitianSpace\right)$ satisfies $\left(\det g\right)^2 = 1$.  In this case, we obtain the connected components of the identity by adding a restriction on the determinant: We have that $G$ consists of all elements $g \in \IsometryGroup\left(\hermitianSpace\right)$ with $\det g = 1$. There are two cases here.
	\begin{enumerate}
		\item Split case: in this case, we can write $$\hermitianSpace = \begin{dcases}
		\xIsotropic \oplus \yIsotropic & \dim_{\finiteField} \hermitianSpace = 2n\\
		\xIsotropic \oplus \finiteField v \oplus \yIsotropic & \dim_{\finiteField}\hermitianSpace = 2n + 1
		\end{dcases}$$ where $\xIsotropic$ and $\yIsotropic$ are totally isotropic (that is, for every $x \in \xIsotropic$, $\innerproduct{x}{x} = 0$ and similarly for $\yIsotropic$) and these spaces are in duality with respect to $\innerproduct{\cdot}{\cdot}$, i.e., the map $\yIsotropic \to \Hom_{\finiteField}\left(\xIsotropic, \finiteField\right)$ given by $y \mapsto \left(x \mapsto \innerproduct{x}{y}\right)$ is an isomorphism. In the odd dimensional case, $v \in \hermitianSpace$ is an element orthogonal to $\xIsotropic$ and $\yIsotropic$ that satisfies $\innerproduct{v}{v} \ne 0$.

@@

	\begin{theorem}
		A maximal rational torus $\algebraicGroup{T}$ of $\algebraicGroup{G} = \algebraicGroup{\UnitaryGroup}_{n}, \algebraicGroup{\Sp}_{2n}, \algebraicGroup{\SO}_{2n + 1}, \algebraicGroup{\SO}_{2n}^{\pm}$ is isomorphic to 	$$\algebraicGroup{T} \cong \restrictionOfScalars{\finiteFieldExtension{\lambda^{+}}}{\finiteField}{\multiplcativeScheme} \times \restrictionOfScalars{\finiteFieldExtension{\lambda^{-}}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1},$$
		where $\left(\lambda^+, \lambda^-\right)$ is a pair of partitions such that $\sizeof{\lambda^+} + \sizeof{\lambda^-} = n$, where in the unitary group case we require that $\lambda^+$ consists only of even parts and that $\lambda^-$ consists only of odd parts. We have that $$\algebraicGroup{T}^{\Frobenius} \cong \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \times \NormOneGroup{2 \lambda^{-}}.$$ The embedding $\algebraicGroup{T}^{\Frobenius} \to \algebraicGroup{G}^{\Frobenius}$ is of the form $$\left(x, y\right) \mapsto
		\begin{pmatrix}
			\iota_+\left(x\right)\\
			& \iota_{-}\left(y\right) &\\
			 & & \involution{\iota_+^{\ast}\left(x\right)}
		\end{pmatrix},$$
		where $\iota_+ \colon \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \to \GL_{\frac{\sizeof{\lambda^+}}{\grpIndex{\quadraticExtension}{\finiteField}}}\left(\quadraticExtension\right)$ and $\iota_- \colon \NormOneGroup{2 \lambda^{-}} \to G'$ are embeddings. Here, $G'$ is a group of the same type as $\algebraicGroup{G}^{\Frobenius}$, with $n$ replaced by $\sizeof{\lambda^{-}}$.

@@

	We summarize these cases in the following theorem.
	\begin{theorem}
	A maximal rational torus $\algebraicGroup{T}$ of $\algebraicGroup{G} =  \algebraicGroup{\GSp}_{2n}, \algebraicGroup{\GSO}_{2n}^{\pm}$ is isomorphic to \begin{align*}
				^{L^{-1}w}\algebraicGroup{T}_0 \cong& \left\{\left(x,y,z\right) \in  \restrictionOfScalars{\finiteFieldExtension{\lambda^+}}{\finiteField}{\multiplcativeScheme} \times \restrictionOfScalars{\finiteFieldExtension{2\lambda^-}}{\finiteField}{\multiplcativeScheme}  \times \multiplcativeScheme \mid \FieldNorm{2\lambda^-}{\lambda^-}(y) = z \right\}.
	\end{align*}
	where $\left(\lambda^+, \lambda^-\right)$ is a pair of partitions such that $\sizeof{\lambda^+} + \sizeof{\lambda^-} = n$. We have that $$\algebraicGroup{T}^{\Frobenius} \cong \left\{\left(x, y, z\right) \in \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \times \multiplicativegroup{\finiteFieldExtension{2 \lambda^-}} \times \multiplicativegroup{\finiteField} \mid \FieldNorm{2 \lambda^{-}}{\lambda^{-}}\left(y\right) = z \right\}.$$ The embedding $\algebraicGroup{T}^{\Frobenius} \to \algebraicGroup{G}^{\Frobenius}$ is of the form $$\left(x, y, z\right) \mapsto
	\begin{pmatrix}
		z \cdot \iota_+ \left(x\right)\\
		& \iota_{-}\left(y\right) &\\
		& & \iota_+^{\ast}\left(x\right)
	\end{pmatrix},$$
	where $\iota_+ \colon \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \to \GL_{\sizeof{\lambda^+}}\left(\finiteField\right)$ and $$\iota_- \colon \left\{ \left(y, z\right)\in \multiplicativegroup{\finiteFieldExtension{2 \lambda^-}} \times \multiplicativegroup{\finiteField} \mid \FieldNorm{2 \lambda^{-}}{\lambda^{-}}\left(y\right) = z \text{ for all } i \right\} \to G'$$ are embeddings. Here, $G'$ is a group of the same type as $\algebraicGroup{G}^{\Frobenius}$, with $n$ replaced by $\sizeof{\lambda^{-}}$.
\end{theorem}
